\chapter{Project Timeline and Work Plan}

This chapter presents the planned timeline of the project, including a detailed work plan and corresponding Gantt charts. The project is divided into two main phases: Phase 1 (September--December) and Phase 2 (January--May).

% =========================
% PHASE 1 – TEXTUAL PLAN
% =========================
\section{Phase 1: September -- December}

\subsection{Kick-off and Background Research (Weeks 1--2)}
\textbf{Main objective:} Establish project foundations and research background.
\begin{itemize}
    \item Environment setup and repository initialization.
    \item Collection of initial lecture materials (videos and slides).
    \item Literature review on RAG systems, sparse/dense/hybrid retrieval, and multimodal embeddings.
    \item Review of RAG frameworks (LangChain, LlamaIndex) and prompt engineering.
    \item Survey of VLMs, OCR tools, and ASR systems.
\end{itemize}

\subsection{Initial Data Processing and Framework Experiments (Weeks 3--4)}
\textbf{Main objective:} Build the initial data pipeline and baseline retrievers.
\begin{itemize}
    \item Audio extraction from lecture videos and ASR experiments.
    \item OCR trials on lecture slides.
    \item Implementation of BM25 and dense retrievers.
    \item Prototype RAG applications using LangChain and LlamaIndex.
    \item Comparative analysis for framework selection.
\end{itemize}

\subsection{Pipeline v1 and Baseline Deployment (Weeks 5--7)}
\textbf{Main objective:} Construct the first end-to-end retrieval pipeline.
\begin{itemize}
    \item Unified evaluation script for BM25, dense, and hybrid retrieval.
    \item Reranker experiments and multi-embedding exploration.
    \item OCR and ASR option benchmarking.
    \item Development of a multi-format document processor.
    \item Initial system architecture design and documentation.
\end{itemize}

\subsection{Pipeline Finalization and Report Drafting (Weeks 8--10)}
\textbf{Main objective:} Stabilize the pipeline and prepare thesis drafts.
\begin{itemize}
    \item Refinement of data preprocessing and embedding pipelines.
    \item Conversion of notebooks into a structured code repository.
    \item Creation of a small evaluation dataset and baseline evaluation.
    \item UI wireframe design and initial backend API planning.
    \item Writing of preliminary report sections.
\end{itemize}

\subsection{Finalization and Presentation Preparation (Weeks 11--12)}
\textbf{Main objective:} Complete Phase 1 deliverables.
\begin{itemize}
    \item Finalization of retrieval experiments and analyses.
    \item Completion of system architecture documentation.
    \item Final report submission and presentation slide preparation.
\end{itemize}

% =========================
% PHASE 1 – GANTT
% =========================
\begin{figure}[H]
\centering
\begin{ganttchart}[
    hgrid,
    vgrid,
    x unit=0.75cm,
    y unit chart=0.6cm,
    bar height=0.55,
    bar label font=\small,
    title label font=\small
]{1}{12}

\gantttitle{Phase 1 Timeline (Weeks)}{12} \\
\gantttitlelist{1,...,12}{1} \\

\ganttbar{Kick-off and Background Research}{1}{2} \\
\ganttbar{Initial Data Processing and Framework Experiments}{3}{4} \\
\ganttbar{Pipeline v1 and Baseline Deployment}{5}{7} \\
\ganttbar{Pipeline Finalization and Report Drafting}{8}{10} \\
\ganttbar{Finalization and Presentation Preparation}{11}{12} \\

\end{ganttchart}
\caption{Phase 1 Timeline(September - December)}
\end{figure}

% =========================
% PHASE 2 – TEXTUAL PLAN
% =========================
\section{Phase 2: January -- May}

\subsection{Phase 2 Kick-off and Integration Planning (Week 13)}
\textbf{Main objective:} Prepare for large-scale integration.
\begin{itemize}
    \item Planning data scaling to 50+ lectures.
    \item Research on video--slide synchronization techniques.
    \item Design of hybrid retrieval architecture with multimodal embeddings.
    \item Backend and frontend sprint planning.
\end{itemize}

\subsection{Core Module Implementation and Dataset Construction (Week 14)}
\textbf{Main objective:} Build multimodal core components.
\begin{itemize}
    \item Implementation of a video--slide synchronization prototype.
    \item Construction of multimodal embedding and vector indexing pipeline.
    \item Collaborative creation of a large-scale evaluation dataset.
\end{itemize}

\subsection{Hybrid RAG Integration and Frontend Development (Weeks 15--17)}
\textbf{Main objective:} Integrate retrieval, backend, and frontend components.
\begin{itemize}
    \item Integration of sparse, dense, and multimodal retrievers using RRF.
    \item Delivery of a unified retrieval function.
    \item Backend--frontend connection and UI component development.
\end{itemize}

\subsection{System Evaluation and Thesis Writing (Weeks 18--20)}
\textbf{Main objective:} Evaluate the system and write core thesis chapters.
\begin{itemize}
    \item Full benchmark evaluation and metric analysis.
    \item Experiments on faithfulness and grounded generation.
    \item Writing of thesis chapters on data pipeline, retrieval, and generation.
\end{itemize}

\subsection{Final Report and Presentation Preparation (Weeks 21--23)}
\textbf{Main objective:} Finalize the thesis and demo system.
\begin{itemize}
    \item Finalization of data processing and retrieval scripts.
    \item UI/UX refinement and demo preparation.
    \item Completion of thesis draft and presentation slides.
\end{itemize}

\subsection{Defense Preparation and Thesis Defense (Weeks 24--25)}
\textbf{Main objective:} Prepare for and complete the thesis defense.
\begin{itemize}
    \item Presentation rehearsals and Q\&A preparation.
    \item Final thesis defense.
\end{itemize}

% =========================
% PHASE 2 – GANTT
% =========================
\begin{figure}[H]
\centering
\begin{ganttchart}[
    hgrid,
    vgrid,
    x unit=0.6cm,
    y unit chart=0.6cm,
    bar height=0.55,
    bar label font=\small,
    title label font=\small
]{1}{13}

\gantttitle{Phase 2 Timeline (Weeks)}{13} \\
\gantttitlelist{13,...,25}{1} \\

\ganttbar{Phase 2 Kick-off and Integration Planning}{1}{1} \\
\ganttbar{Core Module Implementation and Dataset Construction}{2}{2} \\
\ganttbar{Hybrid RAG Integration and Frontend Development}{3}{5} \\
\ganttbar{System Evaluation and Thesis Writing}{6}{8} \\
\ganttbar{Final Report and Presentation Preparation}{9}{11} \\
\ganttbar{Defense Preparation and Thesis Defense}{12}{13} \\

\end{ganttchart}
\caption{Phase 2 Plan (January - May)}
\end{figure}
