\chapter*{Declaration of Authorship}
\addcontentsline{toc}{chapter}{Declaration of Authorship}

We hereby declare that the entire content of this report is the result of research, analysis, evaluation, and implementation conducted by ourselves under the guidance of \textbf{Dr. Nguyen An Khuong} and \textbf{Mr. Nguyen Trang Sy Lam}. The results achieved in this report are authentic and have never been published in any form before. All content presented in this report has been collected, selected, analyzed, and evaluated with rigor throughout the research process. All materials, data, and content referenced from other sources have been properly cited and listed in accordance with regulations in the bibliography.

During the implementation of this report, we have used certain modern support tools, including analysis software, simulation tools, and artificial intelligence (AI) tools, solely for technical support purposes such as language error checking, concept presentation suggestions, or text formatting assistance. All academic content, scientific reasoning, model design, result analysis, and conclusions have been directly conducted by ourselves, ensuring compliance with academic integrity regulations and no violation of plagiarism rules.

While each team member specialized in different aspects of the project: \textbf{Nguyen Quang Phu} concentrated on RAG pipelines and document processing, \textbf{Nguyen Ngoc Khoi} focused on retrieval types and embedding methods, and \textbf{Nguyen Minh Khoi} specialized in Vision-Language Models (VLM), Optical Character Recognition (OCR), and Automatic Speech Recognition (ASR), the team ensured continuous knowledge sharing and collaboration throughout the research process. All members actively participated in discussions, code reviews, and cross-domain learning to maintain a comprehensive understanding of the entire system.

We take full responsibility for this statement of commitment as well as the entire content of this report. If any academic misconduct or improper copying is detected, we accept all disciplinary actions according to the regulations of the Faculty Board and the University Board of Ho Chi Minh City University of Technology - Vietnam National University.

\begin{flushright}
\textit{Ho Chi Minh City, December 2025}
\end{flushright}

\newpage
