\chapter*{Declaration of Authorship}
\addcontentsline{toc}{chapter}{Declaration of Authorship}

We hereby declare that the entire content of this report is the result of research, analysis, evaluation, and implementation conducted by ourselves under the guidance of \textbf{Dr. Nguyen An Khuong} and \textbf{Mr. Nguyen Trang Sy Lam}. The results achieved in this report are authentic and have never been published in any form before. All content presented in this report has been collected, selected, analyzed, and evaluated with rigor throughout the research process. All materials, data, and content referenced from other sources have been properly cited and listed in accordance with regulations in the bibliography.

During the implementation of this report, we have used certain modern support tools, including analysis software, simulation tools, and artificial intelligence (AI) tools, solely for technical support purposes. Specifically, Large Language Model (LLM)-based tools were employed to assist with language refinement, summarization of complex concepts, and drafting support to improve clarity and presentation quality. These tools served as assistive technologies for writing refinement and did not generate the core academic content, scientific reasoning, or technical implementations.

All substantive academic content, scientific reasoning, model design, implementation architecture, result analysis, and conclusions have been directly conducted and authored by ourselves. The use of AI tools in no way substitutes for or replaces the intellectual work, critical thinking, and original contributions that form the foundation of this research. We affirm full responsibility for the accuracy and integrity of all academic claims and technical solutions presented. This disclosure is made in full compliance with contemporary standards for responsible AI usage in academic work and to maintain the highest standards of academic integrity.

While each team member specialized in different aspects of the project, the team ensured continuous knowledge sharing and collaboration throughout the research process. All members actively participated in discussions, code reviews, and cross-domain learning to maintain a comprehensive understanding of the entire system. Details about each member's contribution is mentioned in the appendix.

We take full responsibility for this statement of commitment as well as the entire content of this report. If any academic misconduct or improper copying is detected, we accept all disciplinary actions according to the regulations of the Faculty Board and the University Board of Ho Chi Minh City University of Technology - Vietnam National University.

\begin{flushright}
\textit{Ho Chi Minh City, May 2026}
\end{flushright}

\newpage
